### 2020-06-05 Report/Sections/fourier.tex

% !TeX root = ../main.tex

\section{Model Formulation}

\subsection{Short-time Fourier Transform on Fine-scale process}

One issue with HMMs is that they assume markovian dynamics conditioned on the hidden state, i.e. that any observation $Y^*_t$ depends only on the behaviorial state $X^*_t$ (here we focus on the fine-scale process). However, there are many animal movement processes which violate this markov propertey on very fine scales. For example swimming behaviour of marine mammals can be periodic since the animal repeatedly flukes to propel itself forward. Work has been done in the past to model non-markovian dynamics in the \textit{behavioural} process $X^*_t$ \cite{Langrock:2012}, but addressing non-markovian dynamics within the observation process $Y^*_t$ is still a relatively unstudied area. With improvements in tagging technology allowing for data collection at very high frequencies, data exhibiting noisy and non-markovian fine scale behavior is likely to persist in future studies.

To address this issue, we recommend borrowing techniques from the signal processing literature to compress the data and summarize its essential elements. In particular, we suggest performing the discrete-time Short-time Fourier Transform (STFT) over each observed fine-scale process $Y^*_t$:
%
$$STFT\{Y^*_{t,t^*:t^*+w-1}\}(n) := \hat{Y}^{*(n)}_{t,t^*} = \sum_{n = 0}^{w-1} Y^*_{t,t^*+n}e^{-i \frac{2\pi k}{w} n} \qquad \forall n \in \{0, \ldots, w-1\}, \quad t^* \in \{1,w+1,2w+1, \ldots, w \left(\lfloor T^*_t / w \rfloor -1 \right) + 1\}.$$
%
The STFT slides a moving window of length $w$ accross the time series $Y_t^*$ with a step size of $w$ and transforms the domain of each window from time to frequency. This allows the spectrum of $Y_t^*$ at time $t^*$ to be summerized by a $w$-dimensional vector of fourier coeficients. While other step sizes can be used for the sliding window, we select $w$ to avoid serial dependence between windows.

If $Y^*_t \in \mathbb{R}^{T^*_t}$, then $\hat{Y}_t^* \in \mathbb{C}^{\lfloor T^*_t / w \rfloor \times w}$. While this allows $Y^*_t$ to be represneted in a way that eliminates obvious periodic behavior, the dataset itself is still approximately as large as $Y^*_t$ itself. To reduce the size of $\hat{Y}^*_t$, we propose taking summary statistics of each window as follows:
%
$$Z_{t,t^*}^{*(1)} = \mathcal{R}\left(\hat{Y}^{*(0)}_{t,t^*}\right) \qquad Z_{t,t^*}^{*(2)} = \frac{1}{w}\sum_{n=1}^{\tilde{w}}|\hat{Y}^{*(n)}_{t,t^*}|^2$$
%
$Z_{t,t^*}^{*(1)}$ is equal to the average value of $Y_{t,t^*:t^*+w-1}^*$ while $Z_{t,t^*}^{*(2)}$ is equal to the squared 2-norm of $Y_{t,t^*:t^*+w-1}^*$ that can be attributed to frequencies in the signal between $1$ and $\tilde{w}$ periods per window length. Both the window length $w$ and the max frequency $\tilde{w}$ are tuning parameters that should be tuned in a problem-specific way. $w$ should be long enough to capture the periodic behavior of the underlying process (at least as long as the length of a period), but short enough to avoid over-smoothing of the data and to maintain high resultion in the behavioral process $X^*$. $\tilde{w}$ should be selected such that the maximum frequency of $Y_t^*$ that makes biological sense is $\tilde{w}$ per window length. Note that these summary statistics are just one possible choice, and future studies can adjust the definitions as needed. A visualization of transforming a one-dimensional sequence $Y^*$ to $Z^*$ can be seen in figure (\ref{fig:fourier_example}).

\begin{figure}[h!]
	\centering
	\includegraphics[width=7.5in]{../../Plots/fourier_transform.png}
	\caption{Visualization of transforming $Y^*_t$ into $Z^*_t$ using a sliding window and fourier transform.}
	\label{fig:fourier_example}
\end{figure}

Finally, note that it is possible to accomodate for unequal time steps within each window by using the \textbf{non-uniform discrete Fourier transform (NDFT)}. We do not describe the details of this method in this work, but the generalization is straightforward. Refer to Bagchi et al \cite{Bagchi:2001} for details.

\subsection{Model Structure: combining the HHMM and CarHMM}

Hierarchical hidden Markov models can be used to jointly model simultaneous coarse-scale and fine-scale processes taking place simultaneously. However, as mentioned before, the fine-scale process $Y^*$ can often exhibit autocorrelation and intricate structure. Transforming $Y^*_t$ to $Z^*_t$ removes fine-scale periodic behavior, but $Z^*_t$ can still exhibit autocorrelation, especially in the window averge, $Z_{t,t^*}^{*(1)}$. Therefore, we replace the fine-scale HMM within the Hierarchical HMM with a CarHMM according to figure (\ref{fig:CarHHMM}).

\begin{figure}[h!]
	\centering
	\includegraphics[width=7.5in]{../../Plots/CarHHMM.png}
	\caption{Graphical representation of a CarHHMM. The additional arrows representing autocorrelation between observations are shown in red for emphasis.}
	\label{fig:CarHHMM}
\end{figure}

The likelihood of this model is still easy to calculate using the forward algorithm:
%
$$\calL_{\text{CarHHMM}}(y,z^*;\theta,\theta^*,\Gamma,\Gamma^*,\delta,\delta^*) = \delta P(y_1,z_1^*;\theta,\theta^*,\Gamma^*,\delta^*) \prod_{t=2}^T \Gamma P(y_t,z_t^*;\theta,\theta^*,\Gamma^*,\delta^*) \mathbf{1}$$
%
where:
%
\begin{align*}
P(y_t,z_t^*;\theta,\theta^*,\Gamma^*,\delta^*)  = \text{diag}\left[p_{\theta}(y_t|X_t = x_1)\calL_{\text{CarHMM}}\left(z_t^*;\theta^{*(x_1)},\Gamma^{*(x_1)},\delta^{*(x_1)}\right), . . . , \right.\\
\left. p_{\theta}(y_t|x_t = x_N )\calL_{\text{CarHMM}}\left(z_t^*;\theta^{*(x_N)},\Gamma^{*(x_N)},\delta^{*(x_N)}\right) \right]
\end{align*}
%