### 2020-06-25 Report/Sections/tools.tex

% !TeX root = ../main.tex

\section{Additional Tools}

\subsection{Lag Plot}

One tool that Lawler et al use for data interpretation and model selection is a lag plot. A lag plot is a special case of a scatter plot where the $x$ coordinate of each point corresponds to a datum in a time series (except for the last one), and the $y$ coordinate corresponds to the datum directly after it. In the case of the CarHMM, $(d_t,d_{t+1})$ is plotted for every $t \in \{1, \ldots T-2\}$. If there is no auto-correlation within a given behavioral state, the lag plot should look like a group of spherically symmetric droplet patterns along the line $y=x$, and the number of droplets should correspond to the number of behavioral states. However, if there is significant auto-correlation within states, the droplets will be ``smeared" along the $y=x$ line and the number of behavioral states will be less clear. The lag plot is a good diagnostic tool to visually determine the number of behavioral states in an HMM and whether it is necessary to model within-state autocorrelation using the CarHMM. A lag plot for the killer whale case study is shown in figure (\ref{fig:lag}).

\subsection{Data Interpolation}

As mentioned earlier, the use of an HMM or CarHMM requires observations that are equispaced in time. However, this requirement in not always satisfied for real-world data, especially in the case of marine mammals, where the location of an animal is often only known when that animal comes to the surface. Lawler et al suggest linear interpolation as a solution to this issue, but warn that the uniform time step used for interpolation must be chosen with care. If the time step is too long, then the data will be over-smoothed and important information will be lost from the animal's dynamics. The number of effective observations will also be significantly less than the original data set. If the uniform time step is too short, however, then many interpolated observations will occur on the same line segment, resulting in repetitive step lengths and turning angles equal to zero. See figure (\ref{fig:interpolation}) for a visualization of both issues. To deal with this, Lawler et al. suggest using the following procedure:

\begin{enumerate}
	\item Pick a uniform time step $\Delta t_u$ and interpolate the path of the animal accordingly. Lawler et al suggest that $\Delta t_u$ be somewhere between the median and $3^{rd}$ quartile of all time steps in the raw data set. 
	\item Pick a ``group cutoff level" $\Delta t_g$. Divide observations into groups that are separated by time steps larger than $\Delta t_g$, and treat each group as an independent time series. Lawler et al suggest that $\Delta t_g$ be larger than $\Delta t_u$ to prevent fragmenting the observations too much, but less than $2*\Delta t_u$ to avoid having many interpolated points lie on the same line segment between true observations. Figure (\ref{fig:interpolation_group}) visualizes the issues of selecting $\Delta t_g$ to be too small or too large.
\end{enumerate}

\begin{figure}[h!]
	\centering
	\includegraphics[height=3in]{../Plots/Interpolation.png}
	\caption{Consequences of picking $\Delta t_u$ too large (left) or too small (right). The observed track is shown in black and the interpolated track is shown in red. If $\Delta t_u$ is too large, then the true path will be over-smoothed and information about the animal's path will be lost. If $\Delta t_u$ is too small, then multiple interpolated points will appear on the same line segment, resulting in repeated step sizes and zero turning angles.}
	\label{fig:interpolation}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[height=3in]{../Plots/Interpolation_Group.png}
	\caption{Consequences of picking $\Delta t_g$ too large (above) or too small (below). The observed track is shown in black and the interpolated track is shown in red. If $\Delta t_g$ is too large, then many interpolated points will appear on some line segments, resulting in repeated step lengths and turning angles of zero degrees. If $\Delta t_g$ is too small, then many groups will appear and interpolated points not be able to create paths with turning angles.}
	\label{fig:interpolation_group}
\end{figure}

Even with the above restrictions, there is a considerable amount of freedom in choosing $\Delta t_u$ and $\Delta t_g$. Therefore, Lawler et al. introduce the two statistics to measure the quality of the choice of $\Delta t_u$ and $\Delta t_g$:

$$n_{prop} = \frac{\text{\# of interpolated points}}{\text{\# of observed locations}}$$

$$n_{adj} = \frac{\text{\# of interpolated points} - 2 \cdot \text{\# of groups}}{\text{\# of observed locations} - 2}$$

$n_{prop}$ measures the ratio of interpolated points to observed data points, while $n_{adj}$ measures the ratio of interpolated points that appear in the likelihood calculation to the number of observations that \textit{would} appear in the likelihood calculation had the data not been interpolated. Twice the number of groups is subtracted from the numerator and denominator of $n_{adj}$ because the first and last points of a group have no turning angle associated with them. Lawler et al claim that $n_{prop}$ and $n_{adj}$ should both be as close to one as possible. If these values are much larger than one, the uniform step size may be too small, resulting in replicated (or at least highly correlated) step lengths and turning angles close to zero. If these statistics are much less than one, the uniform step size may be too large, resulting in over-smoothing of the data and lost information.

Lawler et al comment that auto-correlation within the step length sequence is induced by interpolation but claim that further investigation is needed to discover the cause. To establish intuition behind this auto-correlation, suppose an interpolated point $\bfY'_s$ is equal to $\psi_s \bfY_t + (1-\psi_s) \bfY_{t+1}$, another interpolated point $\bfY'_{s+1}$ is equal to $\psi_{s+1} \bfY_{t+1} + (1-\psi_{s+1}) \bfY_{t+2}$, and a third interpolated point $\bfY'_{s+2}$ is equal to $\psi_{s+2} \bfY_{t+2} + (1-\psi_{s+2}) \bfY_{t+3}$, where $0 \leq \psi_s,\psi_{s+1},\psi_{s+2} \leq 1$. This must be the case since a linearly interpolated point is simply a convex combination of two observations. Then, we have:
%
\begin{align*}
	d'_s &= ||\bfY'_{s+1} - \bfY'_s|| \\
	%
	&= ||\psi_{s+1} \bfY_{t+1} + (1-\psi_{s+1}) \bfY_{t+2} - \psi_s \bfY_t - (1-\psi_s) \bfY_{t+1} || \\
	&= ||\psi_{s+1}(\bfY_{t+1} - \bfY_{t+2}) + \psi_s(\bfY_{t+1} - \bfY_t) + \bfY_{t+2} - \bfY_{t+1} ||  \\
	%
	&= ||(1-\psi_{s+1})(\bfY_{t+2} - \bfY_{t+1}) + \psi_s(\bfY_{t+1} - \bfY_t)||
\end{align*}
%
Similarly:
%
$$d'_{s+1} = ||(1-\psi_{s+2})(\bfY_{t+3} - \bfY_{t+2}) + \psi_{s+1}(\bfY_{t+2} - \bfY_{t+1})||$$
%
Both $(1-\psi_{s+1})$ and $\psi_{s+1}$ are positive, and the term $\bfY_{t+2} - \bfY_{t+1}$ shows up in both $d'_s$ and $d'_{s+1}$, so intuitively $d'_s$ and $d'_{s+1}$ should co-vary with one another in most cases. This provides evidence that, in general, the sequence of interpolated step-lengths $\bfD'$ exhibits more auto-correlation than the observed step lengths $\bfD$. It is therefore useful to model auto-correlation using the CarHMM if a uniform time grid is used to interpolate observations.

\subsection{Step-Size Normalization}

Lawler et al recommend dividing all step lengths by the mean observed step length to eliminate units and allow for easy comparison across data sources and animals. This process should be done with care, however, since scaling a random variable can sometimes change its distribution. Luckily, step length in the CarHMM is modeled using a gamma distribution, and gamma random variables still follow a gamma distribution when scaled by a constant. Step-size standardization can only be done if the analysis is done offline, since the mean step length is in flux for online analysis.

%%%%%%%%%%%% THROWAWAY STUFF FOR LATER HERE %%%%%%%%%%

\iffalse

\section{for simulation results}

\subsection{Model Misspecification}

There are two ways to simulate the fourier modes of the acceleration vector $\bfA_{s,t}$:

\begin{enumerate}
	\item $$\bbE_1\left(\bfA_{s,t} | \bfA_{s-1,t}, X^*_{s,t}\right) = \phi \bfA_{s-1,t} + (1-\phi) \mu_{X^*_{s,t}} $$

	\item $$\bbE_2(\bfA_{s,t} | \bfA_{s-1,t}, X^*_{s,t}, X^*_{s-1,t}) = \left\{\begin{array}{lr}
	\phi \bfA_{s-1,t} + (1-\phi) \mu_{X^*_{s,t}} & \text{for } X^*_{s,t} = X^*_{s-1,t}\\
	\mu_{X^*_{s,t}} & \text{for } X^*_{s,t} \neq X^*_{s-1,t}
	\end{array}\right\}$$
\end{enumerate}

\subsubsection{$\bbE_1$}

First, look at $\bbE_1$. This comes with an issue in that when $X^*_{s,t}$ changes values, it will take a while for the process to ``converge" long-run mean $\mu_{X^*_{s,t}}$. In particular, if the behavioral process $X^*$ had been in state 1 for a while and then switches to state 2 for a while, then:

\begin{align*}
	\bbE_1\left(\bfA_{s+1,t}|X^*_{s:s+1,t} = 2, X^*_{s,t} = 1\right) &=  \phi \mu_1 + (1-\phi) \mu_2 \\
	%
	\bbE_1\left(\bfA_{s+2,t}|X^*_{s:s+2,t} = 2, X^*_{s,t} = 1\right) &= \phi\left( \phi \mu_1 + (1-\phi) \mu_2\right) + (1-\phi)\mu_2 \\
	&= \phi^2 \mu_1 + (1-\phi^2) \mu_2\\
	%
	\bbE_1\left(\bfA_{s+n,t}|X^*_{s:s+n,t} = 2, X^*_{s,t} = 1\right) &=  \phi^n \mu_1 + (1-\phi^n) \mu_2 \\
\end{align*}

So, if using $\bbE_1$, there is an assumption that the fourier modes of the acceleration vector $\bfA$ linearly ``converge" to the new value $\mu_i$ from $\mu_j$ when $X^*_{s,t}$ changes from $i$ to $j$. An example of this can be seen in figure (\ref{fig:FOVEDBA}).

 \begin{figure}[h!]
 	\centering
 	\includegraphics[height=3in]{../Code/test.png}
 	\caption{The FoVeDBA of simulated data using $\bbE_1$. Note that the FoVeDBA approaches $\approx$ 300 in state 2 from $\approx$ 10 in state 1 with roughly linear convergence.}
 	\label{fig:FOVEDBA}
 \end{figure}

 Note that if the generative model is correlated, but the inference model isn't, then quick jumps between states 1 and 2 can be missed because the chain never ``converges" to the new mean $\mu_i$, as shown in figure (\ref{fig:STATE}).

\begin{figure}[h!]
	\centering
	\includegraphics[height=3in]{../Code/test2.png}
	\caption{The probability of being in each behavioral state $X^*_{s,t}$ as estimated by the HHMM. The colors represent the true behavioral state.}
	\label{fig:STATE}
\end{figure}

 \subsubsection{$\bbE_2$}

Clearly, the model has issues decodeing the behavioral state when the model is misspecified. Therefore, if the true generative process does not have correlation when jumping between states, then it is better to use $\bbE_2$:

$$\bbE_2(\bfA_{s,t} | \bfA_{s-1,t}, X^*_{s,t}, X^*_{s-1,t}) = \left\{\begin{array}{lr}
\phi \bfA_{s-1,t} + (1-\phi) \mu_{X^*_{s,t}} & \text{for } X^*_{s,t} = X^*_{s-1,t}\\
\mu_{X^*_{s,t}} & \text{for } X^*_{s,t} \neq X^*_{s-1,t}
\end{array}\right\}$$

This fixes the issue of having to ``converge" to the new value of $\mu_i$ when jumping to state $i$. However, note that $\bbE_2$ depends upon both $X^*_{s,t}$ and $X^*_{s-1,t}$. This breaks the assumption in the forward algorithm that $\bfA_{s,t}$ depends only on $X^*_{s,t}$. One way to deal with this issues is to introduce a new hidden state $Z^*_{s,t} = (X^*_{s-1,t},X^*_{s,t})$. This fixes the issue, but now $Z^*_{s,t}$ can take one of $s^2$ values instead of $s$. This grows the state space of our model and can make inference harder.

 \fi

%Figure (\ref{fig:smoothing}) also shows the calculated ODBA as a function of time for one a specific dive. The acceleration exhibits sinusoidal behavior at several points in time which cannot be modeled using HMMs straightforwardly. As a result, the data was split into two-second intervals (100 data points each), and each interval was summarized by the Fourier transform of the ODBA and the velocity at the \textit{end} of the time interval. Velocity was taken at the end of the time interval because the behavior of the whale during the time interval has no impact on the velocity of the whale at the beginning of the time interval. Because velocity is highly correlated with the average acceleration within an interval, the mean was subtracted out of each ODBA time series before the Fourier transform was taken. To reduce the dimension of the ODBA Fourier transform, Fourier coefficients corresponding to frequencies less than or equal to 5 hertz were summed and frequencies above 5 hertz were ignored. One nice property of this process is that it reduces the number of data points by a factor of 100, which drastically speeds up parameter estimation. The optimal length of the time interval over which to take the Fourier transform is a tuning parameter that is difficult to find. An ideal time interval should be long enough give useful Fourier coefficients, but short enough to preserve as much information about the vertical velocity of the animal as possible.

%\subsection{Lag Plot}

%In order to find the appropriate HMM to model whale behavior, a lag plot was made for both velocity and the ODBA fourier sum. The results of doing so are shown in figure (\ref{fig:lag}). Unfortunately, the number of behavioral states is not clear from the lag plot, but it is clear that the velocity exhibits a large degree of autocorrelation. While the ODBA Fourier sum also exhibits some autocorrelation, the relationship is less strong, so autocorrelation was not incorporated in the ODBA Fourier sums emission distribution.