### 2020-07-02 Report/Sections/0627NH.tex
Nancy Revisions

I've replaced your Background section with my Models and Parameter Estimation (NH) section, borrowing heavily from your writing.

Here are some comments and explanations.

General things (to always keep in mind - I may add to this list of rules as we go along)
\begin{enumerate}
\item When I recompile in overleaf, I see there are 21 warnings (in the symbol next to the recompile button).  It looks like these are from sections you have written, mostly from problems with newblock.  Please go through these since Latex may be making decisions that you don't like.  It's always good to go through the Latex errors and warnings. 
    \item Don't begin a sentence with a symbol, e.g.~don't write ``$\Theta$ is an important parameter''. 
    It's better to write ``The parameter $\Theta$ is important''.
    \item If you start a section with a subsection, there needs to be some text before the subsection starts.  This text can be a summary of what is in the section.  I see some writers don't follow my rule!  But it bugs me to have a subsection just start up with no explanation of the section.
    \item  Journal articles do not have as many things in math display mode as a thesis or a report.  This is to save space - printing is expensive.  Oh, there is no printing these days, or very little!  But it is the journal style that developed during the print era.  Of course, if you refer to an equation, it is displayed and numbered. A complicated equation (lots of fractions or layers with subscripts, long expressions) should be displayed since it would be hard to read in text mode, where you aren't allowed to use things like frac for fraction. 
    Deciding whether to display or not is always a judgement call.
    In the end, the copy editor will make decisions, but it's good to have the submission somewhat in a journal style.  
    \item Bibliography things (not crucial at this point but must be done before submission):  these issues are common in all stat journals, not just CJS.  Biology journals tend to use numbers and list papers as they appear in the article.  This is not the case in stat journals.
    \begin{enumerate}
        \item 
    Make sure you have caps where you need them - in your bib file,  markov needs to be \{M\}arkov, for instance, to force capitalization.
    \item You have a bib-style that cites in the text via a number, as opposed to via author (data).   CJS uses author (date).  Also,  CJS doesn't use full first name of authors.  See the sample.pdf document that comes with the CJS template.
    \end{enumerate}
    \item From CJS sample.pdf: ``Try to avoid double subscripts, and never use triple subscripts.''  I think you're good -- no triples, but you do have doubles.  We can take a look at those at some point.   This is a common expectation in stat journals.
    \item From CJS sample.pdf: ``Unless central or essential to the ﬂow of the discussion, mathematical arguments should be deferred to the Appendix.''  So we'll see about that theorem you have.  This is getting more common in stat journals - but it's something you need to check.
    \item From CJS sample.pdf: ``When you submit the ﬁnal version of your manuscript in Latex form, please include postscript ﬁles (.ps or .eps) for the ﬁgures, labelling them ﬁg1.ps, ﬁg2.ps, etc.''  That's a direction for the final submission, after review and acceptance. For this submission, you probably just submit one pdf file (I haven't checked).  If so, I'd put each figure and table on a separate page, not merged into text, at the end of the pdf document.
    \item Vector notation - boldface or not?  matrix notation - caligraphy or not?   I find it easier to keep things straight if vectors are in bold, matrices in cal.   It looks like CJS has no specific rules (some journals do) but seems to allow this (I looked at some published articles). I suggest using bold and caligraphy, but using newcommand in case we want to switch the notation.    What do you think?
\end{enumerate}

Things specific to this version.
\begin{enumerate}
    \item I have commented out the section title lines in your subfiles (the ones you input) and placed the section title lines in main.tex.  This way, Marie and I can see what sections you're planning to include.  But we can also see the titles in the input files, so we don't get confused.
    \item You have done a great job with the notation. I see that is not a simple task. 
    \item
    Some of the write-up can be a little cleaner, technically, and I think some of the explanatory material, although nice, can distract from the technical specifications of the model.  It's good to put the explanatory stuff in the introduction.   See my comments in the introductory section (which I've written in main.tex).
    \begin{enumerate}
    \item Typically there is less detail in a journal article than in, say, a comprehensive proposal.  For the journal article, you can assume that the reader either already knows about things like HMMs or can get enough of an idea from what you have said (possibly supplementing knowledge from your references).   But it is useful to lay out the notation, as you have done. So ... there is a balance!
   \item
     In the end, you consider three models: CarHMM, HHMM, CarHHMM.  So these need to be sharply defined, so the reader can glance back and skim. 
     \item fine-scale, finescale, fine scale.  I keep using different ones!  I think if you use it as a noun, then it is fine scale.  ``The Markov chain is on a fine scale.''  I think if you use it as an adjective, it is fine-scale, as in ``that is a fine-scale model''. 
     What do you think?
     \item HMMs or HMM's,  $X_t$s or $X_t$'s??  I don't know, and I am inconsistent.  I think either is acceptable.  Which do you prefer?
\end{enumerate}
\end{enumerate}