### 2020-08-31 Report/Sections/bibliography.tex

% !TeX root = ../main.tex

\begin{thebibliography}{}
% For English spelling, we follow the style of this dictionary
\bibitem[]{bib1}
Barber, K. (Ed.) (2004). {\it The Oxford Canadian Dictionary}, 2nd ed., 
Oxford University Press, Toronto.

% Single-author journal article
\bibitem[]{bib2}
Cleveland, W. (1979). Robust locally-weighted regression and smoothing scatterplots. 
{\it Journal of the American Statistical Association}, 74, 829-836.

% Genest provided the data for our sample table
\bibitem[]{bib3}
Genest, C. (1999). Probability and statistics: A tale of two worlds? 
{\it The Canadian Journal of Statistics}, 27(2), 421--444.

% Multiple-author journal article.  All the author names are separated by commas, and 
% note the uppercase letter after the colon in the title.
\bibitem[]{bib4}
Johnes, J., Taylor, J., \& Francis, B. (1993). The research
performance of \textsc{uk} universities: A statistical analysis of the results
of the 1989 research selectivity exercise. {\it Journal of the Royal
Statistical Society Series A}, 156, 271--286.

% Book in French.  Be careful with accents.
\bibitem[]{bib5}
Kidron, M. \& Segal, R. (1992). {\it Atlas du Nouvel \'Etat
du Monde}. \'Editions Autrement, Paris.

% Book, two authors, second edition
\bibitem[]{bib6}
Little, R.~J.~A. \& Rubin, D.~B. (2002). {\it Statistical Analysis with Missing Data}, 2nd ed., 
John Wiley \& Sons, New York.

% Unpublished paper with website link
\bibitem[]{bib7}
Molina, I. \& Rao, J.~N.~K. (2009). Small area estimation of poverty indicators. Working Paper 09-15, Statistics and Econometric Series 05, Universidad Carlos III de Madrid, \\http://hdl.handle.net/10016/5644.

% Edited book
\bibitem[]{bib8}
Roberts, G., Ren, Q., \& Rao, J.~N.~K. (2009). 
Using marginal mean models for data from longitudinal surveys
with a complex design: Some advances in methods. In Lynn, R. (Ed.) {\it Methodology of Longitudinal Surveys}, 
John Wiley \& Sons, New York, 351-366.

% !TeX root = main.tex

\bibitem[]{Adam:2019}
Adam, T., Griffiths, C., Leos Barajas, V., Meese, E., Lowe, C., Blackwell, P., Righton, D., \& Langrock, R. (2019).
Joint modelling of multi-scale animal movement data using hierarchical hidden Markov models.
{\it Methods in Ecology and Evolution}, 10(9), 1536–1550.

\iffalse

\bibitem[]{Ailliot:2006,
	author = {Pierre Ailliot},
	year = {2006},
	pages = {271-274},
	title = {Some theoretical results on markov-switching autoregressive models with gamma innovations},
	volume = {343},
	journal = {Comptes Rendus Mathematique},
}

@book{Bagchi:1999,
	author = {Bagchi, Sonali and Mitra, Sanjit},
	year = {1999},
	month = {01},
	pages = {},
	title = {The Nonuniform Discrete Fourier Transform},
	publisher = {Kluwer Academic Publishing},
	doi = {10.1007/978-1-4615-1229-5\_7}
}

\bibitem[]{Barajas:2017,
	author = {Leos-Barajas, Vianey and Gangloff, Eric and Adam, Timo and Langrock, Roland and van Beest, Floris and Nabe-Nielsen, Jacob and Morales, Juan},
	year = {2017},
	month = {02},
	title = {Multi-scale Modeling of Animal Movement and General Behavior Data Using Hidden Markov Models with Hierarchical Structures},
	journal = {Journal of Agricultural Biological and Environmental Statistics},
	doi = {10.1007/s13253-017-0282-9}
}

\bibitem[]{Bebbington:2007,
    author = {Bebbington, Mark},
    year = {2007},
    month = {08},
    pages = {921 - 942},
    title = {Identifying volcanic regimes using Hidden Markov Models},
    volume = {171},
    journal = {Geophysical Journal International},
    doi = {10.1111/j.1365-246X.2007.03559.x}
}

\bibitem[]{Borchers:2013,
    author = {Borchers, David and Zucchini, W and Heide-Jørgensen, Mads Peter and Cañadas, Ana and Langrock, R},
    year = {2013},
    month = {07},
    pages = {},
    title = {Using Hidden Markov Models to Deal with Availability Bias on Line Transect Surveys},
    journal = {Biometrics},
    doi = {10.1111/biom.12049}
}

\bibitem[]{Borger:2020,
    author = {Börger, Luca and Bijleveld, Allert and Fayet, Annette and Machovsky-Capuska, Gabriel and Patrick, Samantha and Street, Garrett and Wal, Eric},
    year = {2020},
    month = {01},
    pages = {6-15},
    title = {Biologging Special Feature},
    volume = {89},
    journal = {Journal of Animal Ecology},
    doi = {10.1111/1365-2656.13163}
}

\bibitem[]{Bulla:2012,
    author = {Bulla, Jan and Lagona, Francesco and Maruotti, Antonello and Picone, Marco},
    year = {2012},
    month = {12},
    pages = {544-567},
    title = {A Multivariate Hidden Markov Model for the Identification of Sea Regimes from Incomplete Skewed and Circular Time Series},
    volume = {17},
    journal = {Journal of Agricultural Biological and Environmental Statistics},
    doi = {10.1007/s13253-012-0110-1}
}

\bibitem[]{Bromback:1998,
    author = { Babette A.   Brumback  and  John A.   Rice },
    title = {Smoothing Spline Models for the Analysis of Nested and Crossed Samples of Curves},
    journal = {Journal of the American Statistical Association},
    volume = {93},
    number = {443},
    pages = {961-976},
    year  = {1998},
    publisher = {Taylor & Francis},
    doi = {10.1080/01621459.1998.10473755}
}

\bibitem[]{Cade:2017,
    author = {Cade, David and Barr, Kelly and Calambokidis, John and Friedlaender, Ari and Goldbogen, Jeremy},
    year = {2017},
    month = {11},
    pages = {jeb.170449},
    title = {Determining forward speed from accelerometer jiggle in aquatic nvironments},
    volume = {221},
    journal = {The Journal of Experimental Biology},
    doi = {10.1242/jeb.170449}
}

\bibitem[]{Chen:2012,
	Author = {Kehui Chen and Hans-Georg M{\"u}ller},
	Date-Modified = {2020-07-26 09:26:32 -0700},
	Issn = {01621459},
	Journal = {Journal of the American Statistical Association},
	Number = {500},
	Pages = {1599--1609},
	Publisher = {[American Statistical Association, Taylor & Francis, Ltd.]},
	Title = {Modeling Repeated Functional Observations},
	Url = {http://www.jstor.org/stable/23427358},
	Volume = {107},
	Year = {2012}
}

\bibitem[]{Crainiceanu:2009,
    author = {Crainiceanu, Ciprian and Staicu, Ana-Maria and Di, Chongzhi},
    year = {2009},
    month = {12},
    pages = {1550-1561},
    title = {Generalized Multilevel Functional Regression},
    volume = {104},
    journal = {Journal of the American Statistical Association},
    doi = {10.1198/jasa.2009.tm08564}
}

\bibitem[]{DeRuiter:2017,
    author = {DeRuiter, Stacy and Langrock, Roland and Skirbutas, Tomas and Goldbogen, Jeremy and Calambokidis, John and Friedlaender, Ari and Southall, Brandon},
    year = {2017},
    month = {03},
    pages = {362-392},
    title = {A multivariate mixed hidden Markov model for blue whale behaviour and responses to sound exposure},
    volume = {11},
    journal = {The Annals of Applied Statistics},
    doi = {10.1214/16-AOAS1008}
}

\bibitem[]{Di:2009,
	Author = {Di, Chong-Zhi and Crainiceanu, Ciprian M. and Caffo, Brian S. and Punjabi, Naresh M.},
	Doi = {10.1214/08-AOAS206},
	Fjournal = {Annals of Applied Statistics},
	Journal = {Ann. Appl. Stat.},
	Month = {03},
	Number = {1},
	Pages = {458--488},
	Publisher = {The Institute of Mathematical Statistics},
	Title = {Multilevel functional principal component analysis},
	Url = {https://doi.org/10.1214/08-AOAS206},
	Volume = {3},
	Year = {2009}
	}

\bibitem[]{Douc:2004,
	author = {Douc, Randal and Moulines, Eric and Rydén, Tobias},
	year = {2004},
	month = {10},
	pages = {2254-2304},
	title = {Asymptotic properties of the maximum likelihood estimator in autoregressive models with Markov regime},
	volume = {32},
	journal = {The Annals of Statistics},
	doi = {10.1214/009053604000000021}
}

\bibitem[]{Dot:2016,
    author = {Jeanniard du Dot, Tiphaine and Trites, Andrew and Arnould, John and Speakman, John and Guinet, Christophe},
    year = {2016},
    month = {10},
    pages = {},
    title = {Activity-specific metabolic rates for diving, transiting, and resting at sea can be estimated from time-activity budgets in free-ranging marine mammals},
    volume = {7},
    journal = {Ecology and Evolution},
    doi = {10.1002/ece3.2546}
}

\bibitem[]{Fehlmann:2017,
    author = {Fehlmann, Gaelle and O'Riain, J.M. and Hopkins, Phil and O'Sullivan, Jack and Holton, Mark and Shepard, Emily and King, Andrew},
    year = {2017},
    month = {12},
    pages = {},
    title = {Identification of behaviours from accelerometer data in a wild social primate},
    volume = {5},
    journal = {Animal Biotelemetry},
    doi = {10.1186/s40317-017-0121-3}
}

\bibitem[]{Fischer:1993,
  title={The Markov-Modulated Poisson Process (MMPP) Cookbook},
  author={Wolfgang Fischer and Kathleen S. Meier-Hellstern},
  journal={Perform. Evaluation},
  year={1993},
  volume={18},
  pages={149-171}
}

\bibitem[]{Ford:2006,
    author = {Ford, John and Ellis, Graeme},
    year = {2006},
    month = {07},
    pages = {185-199},
    title = {Selective foraging by fish-eating killer whales Orcinus orca in British Columbia},
    volume = {316},
    journal = {Marine Ecology-progress Series - MAR ECOL-PROGR SER},
    doi = {10.3354/meps316185}
}

@techreport{Ford:2015,
    author = {Ford, John et al.},
    year = {2015},
    month = {12},
    title = {Status review update for Pacific salmon and steelhead listed under the Endangered Species Act: Pacific Northwest},
    institution = {National Marine Fisheries Service, Northwest Fisheries Science Center},
    url = {http://www.westcoast.fisheries.noaa.gov/publications/status_reviews/salmon_steelhead/2016_status_review.html}
}

\bibitem[]{Fu:2017,
    author = {Fu, Eric and Heckman, Nancy},
    year = {2017},
    month = {12},
    pages = {},
    title = {Model-based curve registration via stochastic approximation EM algorithm},
    journal = {Computational Statistics and Data Analysis},
    doi = {10.1016/j.csda.2018.06.010}
}

\bibitem[]{Gao:2012,
	author = {Gao, Fuchang and Han, Lixing},
	year = {2012},
	month = {05},
	pages = {259-277},
	title = {Implementing the Nelder-Mead simplex algorithm with adaptive parameters},
	volume = {51},
	journal = {Computational Optimization and Applications},
	doi = {10.1007/s10589-010-9329-3}
}

@inproceedings{Getman:2009,
    author = {Getman, Anya and Cooper, Christopher and Key, Gary and Zhou, Heng and Frankle, Nick},
    year = {2009},
    month = {05},
    pages = {101 - 104},
    title = {Detection of mobile machine damage using accelerometer data and prognostic health monitoring techniques},
    doi = {10.1109/CIVVS.2009.4938730}
}

\bibitem[]{Gleiss:2011,
	author = {Gleiss, Adrian and Wilson, Rory and Shepard, Emily},
	year = {2011},
	month = {01},
	pages = {23-33},
	title = {Making overall dynamic body acceleration work: On the theory of acceleration as a proxy for energy expenditure},
	journal = {Methods in Ecology and Evolution},
	doi = {10.1111/j.2041-210X.2010.00057.x}
}

\bibitem[]{Green:2009,
	author = {Green, J. A. and Halsey, L. G. and Wilson, R. P. and Frappell, P. B.},
	title = {Estimating energy expenditure of animals using the accelerometry technique: activity, inactivity and comparison with the heart-rate technique},
	volume = {212},
	number = {5},
	pages = {745--746},
	year = {2009},
	doi = {10.1242/jeb.030049},
	publisher = {The Company of Biologists Ltd},
	issn = {0022-0949},
	URL = {https://jeb.biologists.org/content/212/5/745},
	eprint = {https://jeb.biologists.org/content/212/5/745.full.pdf},
	journal = {Journal of Experimental Biology}
}

\bibitem[]{Gruen:2017,
    author = {Gruen, Margaret and Alfaro Córdoba, Marcela and Thomson, Andrea and Worth, Alicia and Staicu, Ana-Maria and Lascelles, B Duncan X},
    year = {2017},
    month = {01},
    pages = {e0169576},
    title = {The Use of Functional Data Analysis to Evaluate Activity in a Spontaneous Model of Degenerative Joint Disease Associated Pain in Cats},
    volume = {12},
    journal = {PLOS ONE},
    doi = {10.1371/journal.pone.0169576}
}

\bibitem[]{Hastie:2006,
    author = {Hastie, Gordon and Rosen, David and Trites, Andrew},
    year = {2006},
    month = {09},
    pages = {163-170},
    title = {The influence of depth on a breath-hold diver: Predicting the diving metabolism of Steller sea lions (Eumetopias jubatus)},
    journal = {Journal of Experimental Marine Biology and Ecology},
    doi = {10.1016/j.jembe.2006.05.004}
}

\bibitem[]{Heerah:2017,
    author = {Heerah, Karine and Woillez, Mathieu and Fablet, Ronan and Garren, François and Martin, Stéphane and De Pontual, Hélène},
    year = {2017},
    month = {09},
    pages = {},
    title = {Coupling spectral analysis and hidden Markov models for the segmentation of behavioural patterns},
    volume = {5},
    journal = {Movement ecology},
    doi = {10.1186/s40462-017-0111-3}
}

\bibitem[]{Hooten:2017,
	author = {Hooten, Mevin and King, Ruth and Langrock, Roland},
	year = {2017},
	month = {08},
	pages = {},
	title = {Guest Editor’s Introduction to the Special Issue on “Animal Movement Modeling”},
	journal = {Journal of Agricultural, Biological and Environmental Statistics},
	doi = {10.1007/s13253-017-0299-0}
}

\bibitem[]{Isojunno:2017,
    author = {Isojunno, Saana and Sadykova, Dina and DeRuiter, S. and Curé, Charlotte and Visser, Fleur and Thomas, Len and Miller, P. and Harris, C.},
    year = {2017},
    month = {12},
    pages = {},
    title = {Individual, ecological, and anthropogenic influences on activity budgets of long‐finned pilot whales},
    volume = {8},
    journal = {Ecosphere},
    doi = {10.1002/ecs2.2044}
}

\bibitem[]{Juang:1991,
    author = { B. H.   Juang  and  L. R.   Rabiner },
    title = {Hidden Markov Models for Speech Recognition},
    journal = {Technometrics},
    volume = {33},
    number = {3},
    pages = {251-272},
    year  = {1991},
    publisher = {Taylor & Francis},
    doi = {10.1080/00401706.1991.10484833}
}

\bibitem[]{Johnson:2008,
	author = {Johnson, Devin and London, Josh and Lea, Mary-Anne and Durban, John},
	year = {2008},
	month = {06},
	pages = {1208-15},
	title = {Continuous-time correlated random walk model for animal telemetry data},
	volume = {89},
	journal = {Ecology},
	doi = {10.1890/07-1032.1}
}

@inbook{Kokoszka:2018,
	Author = {P Kokoszka and M Reimherr},
	Chapter = {8},
	Date-Added = {2020-07-26 09:36:01 -0700},
	Date-Modified = {2020-07-26 09:37:50 -0700},
	Publisher = {Chapman and Hall/CRC},
	Title = {Introduction to Functional Data Analysis},
	Year = {2018}
}

\bibitem[]{Langrock:2012,
	author = {Roland Langrock and Ruth King and Jason Matthiopoulos and Len Thomas and Daniel Fortin and Juan Morales},
	year = {2012},
	month = {11},
	pages = {2336-42},
	title = {Flexible and practical modeling of animal telemetry data: Hidden Markov models and extensions},
	volume = {93},
	journal = {Ecology},
	doi = {10.2307/41739305}
}

\bibitem[]{Langrock:2013,
    author = {Langrock, Roland and Swihart, Bruce and Caffo, Brian and Punjabi, Naresh and Crainiceanu, Ciprian},
    year = {2013},
    month = {08},
    pages = {},
    title = {Combining hidden Markov models for comparing the dynamics of multiple sleep electroencephalograms},
    volume = {32},
    journal = {Statistics in medicine},
    doi = {10.1002/sim.5747}
}

\bibitem[]{Langrock:2018,
	author = {R Langrock and T Adam and V Leos‐Barajas and S Mews S and DL Miller and YP Papastamatiou YP},
	year = {2018},
	month = {},
	pages = {179-200},
	volume = {72},
	title = {Spline‐based nonparametric inference in general state‐switching models},
	journal = {Statistica Neerlandica},
	doi = {10.1111/stan.12133} 
}

\bibitem[]{Lawler:2019,
	author = {Lawler, Ethan and Whoriskey, Kim and Aeberhard, William and Field, Chris and Mills Flemming, Joanna},
	year = {2019},
	month = {05},
	pages = {},
	title = {The Conditionally Autoregressive Hidden Markov Model (CarHMM): Inferring Behavioural States from Animal Tracking Data Exhibiting Conditional Autocorrelation},
	journal = {Journal of Agricultural, Biological and Environmental Statistics},
	doi = {10.1007/s13253-019-00366-2}
}

\bibitem[]{Lewin:2011,
	author = {Lewin-Koh, Nicholas and Bivand, Roger and Pebesma, Edzer and Archer, E and Baddeley, A and Bibiko, Hans-Jörg and Dray, Stéphane and Forrest, D and Friendly, M and Giraudoux, Patrick},
	year = {2011},
	month = {01},
	pages = {},
	title = {maptools: Tools for reading and handling spatial objects. R package version 0.8-27}
}

\bibitem[]{Liu:2015,
    author = {Liu, Yu-Ying and Li, Shuang and Li, Fuxin and Song, Le and Rehg, James},
    year = {2015},
    month = {01},
    pages = {3599-3607},
    title = {Efficient Learning of Continuous-Time Hidden Markov Models for Disease Progression},
    volume = {28},
    journal = {Advances in neural information processing systems}
}

@inproceedings{Lucero:2019,  
    author={P. {Lucero} and R. {Sánchez} and J. {Macancela} and D. {Cabrera} and M. {Cerrada} and C. {Li} and H. R. {Alonso}},  booktitle={2019 Prognostics and System Health Management Conference (PHM-Paris)},  
    title={Accelerometer Placement Comparison for Crack Detection in Railway Axles Using Vibration Signals and Machine Learning},   
    year={2019},  
    volume={},  
    number={},  
    pages={291-296}
}

\bibitem[]{McClintock:2017,
	author = {McClintock, Brett},
	year = {2017},
	month = {06},
	pages = {1-21},
	title = {Incorporating Telemetry Error into Hidden Markov Models of Animal Movement Using Multiple Imputation},
	journal = {Journal of Agricultural, Biological, and Environmental Statistics},
	doi = {10.1007/s13253-017-0285-6}
}

\bibitem[]{McClintock:2020,
    author = {McClintock, Brett and Langrock, Roland and Gimenez, Olivier and Cam, Emmanuelle and Borchers, David and Glennie, Richard and Patterson, Toby},
    year = {2020},
    month = {02},
    pages = {},
    title = {Uncovering ecological state dynamics with hidden Markov models},
    journal = {Preprint}
}

\bibitem[]{Michelot:2019,
	author = {Michelot, Th\`eo and Blackwell, Paul},
	year = {2019},
	month = {01},
	pages = {},
	title = {State-switching continuous-time correlated random walks},
	journal = {Methods in Ecology and Evolution},
	doi = {10.1111/2041-210X.13154}
}

\bibitem[]{Morales:2004,
	author = {Morales, Juan and Haydon, Daniel and Frair, Jacqui and Holsinger, Kent and Fryxell, John},
	year = {2004},
	month = {09},
	pages = {},
	title = {Extracting More out of Relocation Data: Building Movement Models as Mixtures of Random Walks},
	volume = {85},
	journal = {EEB Articles},
	doi = {10.1890/03-0269}
}

\bibitem[]{Morris:2007,
	Author = {Jeffrey S Morris and Cassandra Arroyo and Brent A Coull and Louise M Ryan and Richard Herrick and Steven L Gortmaker},
	Doi = {10.1198/016214506000000465},
	Journal = {Journal of the American Statistical Association},
	Number = {476},
	Pages = {1352-1364},
	Publisher = {Taylor & Francis},
	Title = {Using Wavelet-Based Functional Mixed Models to Characterize Population Heterogeneity in Accelerometer Profiles},
	Volume = {101},
	Year = {2006}
}

\bibitem[]{Noren:2011,
    author = {Noren, Dawn},
    year = {2011},
    month = {01},
    pages = {60 - 77},
    title = {Estimated field metabolic rates and prey requirements of resident killer whales},
    volume = {27},
    journal = {Marine Mammal Science},
    doi = {10.1111/j.1748-7692.2010.00386.x}
}

@misc{Nunes:2018,
	author = {Alex Nunes},
	title = {Divebomb},
	url = {https://github.com/ocean-tracking-network/divebomb/blob/master/docs/index.rst},
	year = {2018},
	month = {07}
}

\bibitem[]{Patterson:2017,
	author = {Patterson, Toby and Parton, Alison and Langrock, Roland and Blackwell, Paul and Thomas, Len and King, Ruth},
	year = {2017},
	month = {07},
	pages = {},
	title = {Statistical modelling of individual animal movement: an overview of key methods and a discussion of practical challenges},
	journal = {AStA Advances in Statistical Analysis},
	doi = {10.1007/s10182-017-0302-7}
}

\bibitem[]{Pohle:2017,
	author = {Pohle, Jennifer and Langrock, Roland and van Beest, Floris and Schmidt, Niels},
	year = {2017},
	month = {06},
	pages = {1-24},
	title = {Selecting the Number of States in Hidden Markov Models: Pragmatic Solutions Illustrated Using Animal Movement},
	volume = {22},
	journal = {Journal of Agricultural, Biological, and Environmental Statistics},
	doi = {10.1007/s13253-017-0283-8}
}

\bibitem[]{Qasem:2012,
    author = {Qasem, Lama and Cardew, Antonia and Wilson, Alexis and Griffiths, Iwan and Halsey, Lewis and Shepard, Emily and Gleiss, Adrian and Wilson, Rory},
    year = {2012},
    month = {02},
    pages = {e31187},
    title = {Tri-Axial Dynamic Acceleration as a Proxy for Animal Energy Expenditure; Should We Be Summing Values or Calculating the Vector?},
    volume = {7},
    journal = {PloS one},
    doi = {10.1371/journal.pone.0031187}
}

@book{Ramsay:2005,
	Author = {J O Ramsay and Bernard Silverman},
	Date-Added = {2020-07-26 09:30:43 -0700},
	Date-Modified = {2020-07-26 09:32:18 -0700},
	Publisher = {Springer},
	Title = {Functional Data Analysis},
	Year = {2005}}

\bibitem[]{Rice:2001,
    author = {Rice, John and Wu, Colin},
    year = {2001},
    month = {04},
    pages = {253-9},
    title = {Nonparametric Mixed Effects Models for Unequally Sampled Noisy Curves},
    volume = {57},
    journal = {Biometrics},
    doi = {10.1111/j.0006-341X.2001.00253.x}
}

@book{Serfozo:2009,
	title = {Basics of Applied Stochastic Processes},
	author = {Serfozo, Richard},
	year = {2009},
	publisher = {Springer-Verlag Berlin Heidelberg},
	doi = {10.1007/978-3-540-89332-5},
	ISSN = {1431-7028}
}

\bibitem[]{Shorter:2017,
    author = {Shorter, Kenneth and Shao, Yunli and Ojeda, Lauro and Barton, Kira and Rocho-Levine, Julie and Van der Hoop, Julie and Moore, Michael},
    year = {2017},
    month = {06},
    pages = {},
    title = {A day in the life of a dolphin: Using bio-logging tags for improved animal health and well-being},
    journal = {Marine Mammal Science},
    doi = {10.1111/mms.12408}
}

\bibitem[]{Simon:2012,
    author = {Simon, Malene and Johnson, Mark and Madsen, Peter},
    year = {2012},
    month = {11},
    pages = {3786-98},
    title = {Keeping momentum with a mouthful of water: Behavior and kinematics of humpback whale lunge feeding},
    volume = {215},
    journal = {The Journal of experimental biology},
    doi = {10.1242/jeb.071092}
}

\bibitem[]{Sutherland:1998,
    author = {Sutherland, William},
    year = {1998},
    month = {11},
    pages = {801-809},
    title = {The importance of behavioral studies in conservation biology},
    volume = {56},
    journal = {Animal Behaviour},
    doi = {10.1006/anbe.1998.0896}
}

\bibitem[]{Tennessen:2019a,
    author = {Tennessen, Jennifer and Holt, Marla and Hanson, M. and Emmons, Candice and Giles, Deborah and Hogan, Jeffrey},
    year = {2019},
    month = {02},
    pages = {},
    title = {Kinematic signatures of prey capture from archival tags reveal sex differences in killer whale foraging activity},
    volume = {222},
    journal = {The Journal of Experimental Biology},
    doi = {10.1242/jeb.191874}
}

\bibitem[]{Tennessen:2019b,
    author = {Tennessen, Jennifer and Holt, Marla and Ward, Eric and Hanson, M. and Emmons, Candice and Giles, Deborah and Hogan, Jeffrey},
    year = {2019},
    month = {10},
    pages = {14951},
    title = {Hidden Markov models reveal temporal patterns and sex differences in killer whale behavior},
    volume = {9},
    journal = {Scientific Reports},
    doi = {10.1038/s41598-019-50942-2}
}

\bibitem[]{Virtanen:2019,
	author = {{Virtanen}, Pauli and {Gommers}, Ralf and {Oliphant},
	Travis E.  and {Haberland}, Matt and {Reddy}, Tyler and
	{Cournapeau}, David and {Burovski}, Evgeni and {Peterson}, Pearu and {Weckesser}, Warren and {Bright}, Jonathan and {van der Walt},
	St{\'e}fan J.  and {Brett}, Matthew and {Wilson}, Joshua and
	{Jarrod Millman}, K.  and {Mayorov}, Nikolay and {Nelson}, Andrew
	R.~J. and {Jones}, Eric and {Kern}, Robert and {Larson}, Eric and
	{Carey}, CJ and {Polat}, {\.I}lhan and {Feng}, Yu and {Moore},
	Eric W. and {Vand erPlas}, Jake and {Laxalde}, Denis and
	{Perktold}, Josef and {Cimrman}, Robert and {Henriksen}, Ian and
	{Quintero}, E.~A. and {Harris}, Charles R and {Archibald}, Anne M.
	and {Ribeiro}, Ant{\^o}nio H. and {Pedregosa}, Fabian and
	{van Mulbregt}, Paul and {Contributors}, SciPy 1. 0},
	title = "{SciPy 1.0--Fundamental Algorithms for Scientific Computing in Python}",
	journal = {arXiv e-prints},
	year = "2019",
	month = "Jul",
	eid = {arXiv:1907.10121},
	pages = {arXiv:1907.10121},
	archivePrefix = {arXiv},
	eprint = {1907.10121},
	primaryClass = {cs.MS},
	adsurl = {https://ui.adsabs.harvard.edu/abs/2019arXiv190710121V},
	adsnote = {Provided by the SAO/NASA Astrophysics Data System}
}

\bibitem[]{Viterbi:1967,
	author = {Viterbi, A.},
	year = {1967},
	month = {April},
	pages = {260-269},
	title = {Error bounds for convolutional codes and an asymptotically optimum decoding algorithm},
	journal = {IEEE Transactions on Information Theory},
	doi = {10.1109/TIT.1967.1054010}
}

\bibitem[]{Whoriskey:2016,
	author = {Whoriskey, Kim and Auger-M\'eth\'e, Marie and Albertsen, Christoffer and Whoriskey, F. and Binder, Thomas and Krueger, Charles and Flemming, Joanna},
	year = {2016},
	month = {12},
	pages = {},
	title = {A Hidden Markov Movement Model for rapidly identifying behavioral states from animal tracks},
	volume = {7},
	journal = {Ecology and Evolution},
	doi = {10.1002/ece3.2795}
}

\bibitem[]{Williams:1999,
    author = {Williams, Terrie and Haun, J and Friedl, WA},
    year = {1999},
    month = {11},
    pages = {2739-48},
    title = {The diving physiology of bottlenose dolphins (Tursiops truncatus): I. Balancing the demands of exercise for energy conservation at depth},
    volume = {202},
    journal = {The Journal of experimental biology}
}

\bibitem[]{Williams:2009,
    author = {Williams, Rob and Noren, Dawn},
    year = {2009},
    month = {04},
    pages = {327 - 350},
    title = {Swimming speed, respiration rate, and estimated cost of transport in adult killer whales},
    volume = {25},
    journal = {Marine Mammal Science},
    doi = {10.1111/j.1748-7692.2008.00255.x}
}

\bibitem[]{Wilson:2019,
    author = {Wilson, Rory and Börger, Luca and Holton, Mark and Scantlebury, D. and Laich, Agustina and Quintana, Flavio and Rosell, Frank and Graf, Patricia and Williams, Hannah and Gunner, Richard and Hopkins, Lloyd and Marks, Nikki and Geraldi, Nathan and Duarte, Carlos and Scott, Rebecca and Strano, Michael and Robotka, Hermina and Eizaguirre, Christophe and Fahlman, Andreas and Shepard, Emily},
    year = {2019},
    month = {06},
    pages = {},
    title = {Estimates for energy expenditure in free-living animals using acceleration proxies: A reappraisal},
    volume = {89},
    journal = {Journal of Animal Ecology},
    doi = {10.1111/1365-2656.13040}
}

\bibitem[]{Wright:2017,
    author = {Wright, Brianna and Ford, John and Ellis, Graeme and Deecke, Volker and Shapiro, Ari and Battaile, Brian and Trites, Andrew},
    year = {2017},
    month = {12},
    pages = {},
    title = {Fine-scale foraging movements by fish-eating killer whales (Orcinus orca) relate to the vertical distributions and escape responses of salmonid prey (Oncorhynchus spp.)},
    volume = {5},
    journal = {Movement Ecology},
    doi = {10.1186/s40462-017-0094-0}
}

\bibitem[]{Xiao:2014,
    author = {Xiao, Luo and Huang, Lei and Schrack, Jennifer and Ferrucci, Luigi and Zipunnikov, Vadim and Crainiceanu, Ciprian},
    year = {2014},
    month = {10},
    pages = {},
    title = {Quantifying the lifetime circadian rhythm of physical activity: A covariate-dependent functional approach},
    volume = {16},
    journal = {Biostatistics (Oxford, England)},
    doi = {10.1093/biostatistics/kxu045}
}

\bibitem[]{Xu:2018,
    author = {Xu, Zekun and Laber, Eric and Staicu, Ana-Maria},
    year = {2018},
    month = {12},
    pages = {},
    title = {Hierarchical Continuous Time Hidden Markov Model, with Application in Zero-Inflated Accelerometer Data},
    journal = {Preprint}
}

\bibitem[]{Yao:2005,
    author = {Yao, Fang and Muller, Hans-Georg and Wang, Jane-Ling},
    year = {2005},
    month = {06},
    pages = {577-590},
    title = {Functional Data Analysis for Sparse Longitudinal Data},
    volume = {100},
    journal = {Journal of the American Statistical Association},
    doi = {10.1198/016214504000001745}
}

@book{Zucchini:2016,
	author = {Zucchini, Walter and Macdonald, Iain and Langrock, Roland},
	year = {2016},
	month = {06},
	publisher = {CRC Press},
	title = {Hidden Markov Models for Time Series - An Introduction Using R},
	isbn = {9781482253832}
}

\fi

\end{thebibliography}