### 2021-05-22 Report/Sections/models.tex

%\section{Models and parameter estimation}

Consider a sequence of $T$ curves, where curve $t$ is characterized by a curve-level (or coarse-scale) observation $Y_t$ as well as a sequence of $T^*_t$ within-curve (or fine-scale) observations $Y^*_{t}$. Namely, $Y^*_{t} \equiv \big\{Y^*_{t,1},\ldots,Y^*_{t,T^*_t}\big\}$ is made up of fine-scale quantities derived from curve $t$ indexed by $t^*$. Both $Y_t$ and $Y^*_{t,t^*}$ can be either vectors or scalars. We call the sequence of coarse-scale observations $Y \equiv \big\{Y_1, \ldots, Y_T\big\}$ and the collection of all fine-scale observations $Y^* \equiv \big\{Y^*_1,\ldots,Y^*_T \big\}$. To develop our model for this data, we detail the structure of a traditional HMM followed by three variations which generalize its base structure. We then show how each of these generalized HMMs can be synthesized to form a wide variety of more complicated models.

\subsection{HMMs as a base structure}
\label{subsec:HMM}

Hidden Markov models describe state-switching Markovian processes in discrete time and are the core structure we use to model both $Y$ and $Y^*$. For simplicity we focus on $Y$ to introduce the model. An HMM is comprised of a sequence of unobserved states $X \equiv \big\{X_1, \ldots, X_T\big\}$ together with an observation sequence $Y \equiv \big\{Y_1, \ldots, Y_T\big\}$, where $X_t$ is associated with the observation $Y_t$. The $Y_t$'s are often referred to as ``emissions'' and the index $t$ typically refers to time. 
The $X_t$'s form a Markov chain and can take integer values between $1$ and $N$. Their distribution is governed by the distribution of the initial state $X_1$ and an $N \times N$ transition probability matrix $\Gamma$, where $\Gamma_{ij} = \Pr(X_{t+1} = j | X_t = i)$. 
%
We assume that $X_1$ follows the chain's stationary distribution, which is denoted by an $N$-dimensional row vector $\delta$, where
$\delta_i = \Pr(X_1 = i).$
A Markov chain's stationary distribution is determined by its probability transition matrix via $\delta = \delta \Gamma$ and $\sum_{i=1}^N \delta_i = 1$.
%
The distribution of an emission $Y_t$ conditioned on the corresponding hidden state $X_t$ does not depend upon any other observation or hidden state.
%
If $X_t=i$ then we denote the conditional density or probability mass function of $Y_t$ as $f^{(i)}(\cdot ; \theta^{(i)})$ or simply $f^{(i)}(\cdot)$, where $\theta^{(i)}$ is a state-dependent parameter describing the emission distribution.
%

Using observation emissions, denoted here as $y \equiv \{y_1,\ldots,y_T\}$, we can find the maximum likelihood estimates of the parameters $\Gamma$ and $\theta \equiv \{\theta^{(1)},\ldots,\theta^{(N)}\}$. The likelihood $\calL_{\text{HMM}}$ can be evaluated using the well-known \textit{forward algorithm} \citep{Zucchini:2016}:

\[\calL_{\text{HMM}}(\theta,\Gamma;y) = \delta P(y_1;\theta) \prod_{t=2}^T \Gamma P(y_t;\theta) \mathbf{1}_N,\]
%
where $\mathbf{1}_N$ is an $N$-dimensional column vector of ones and
%
$P(y_t;\theta)$ is an $N \times N$ diagonal matrix with $(i,i)^{th}$ entry $f^{(i)}(y_t; \theta^{(i)})$.
%

Following \citet{Barajas:2017}, we reparameterize the $N \times N$ transition probability matrix $\Gamma$ such that the entries of the matrix are forced to be non-negative and the rows sum to 1:

\[
\Gamma_{ij} = \frac{\exp(\eta_{ij})}{\sum_{k=1}^N \exp(\eta_{ik})}, 
\]
%
where $i,j = 1,\ldots,N$ and $\eta_{ii}$ is set to zero for identifiability. This simplifies likelihood maximization by removing constraints in the optimization problem. For simplicity, we will continue to use $\Gamma$ in our notation, suppressing the reparameterization in terms of $\eta$. Figure \ref{fig:models}a represents the dependence structure of an HMM.

\subsection{Relaxing conditional independence with the CarHMM}
\label{subsec:CarHMM}

A conditionally auto-regressive hidden Markov model, or CarHMM \citep{Lawler:2019}, is a generalization of an HMM which explicitly models auto-correlation in the observation sequence beyond the correlation induced by the hidden state process. Like a traditional HMM, a CarHMM is made up of a Markov chain of unobserved states $\{X_1,\ldots,X_T\}$ each taking integer values between $1$ and $N$. A CarHMM also has a transition probability matrix $\Gamma$ and initial distribution $\delta$ equal to the stationary distribution of $\Gamma$. Unlike a traditional HMM, the CarHMM assumes that the distribution of $Y_t$ conditioned on $X_1,\ldots, X_T$ and $Y_1,\ldots, Y_{t-1}$ depends on both $X_t$ and $Y_{t-1}$ rather than only $X_t$. The first emission $Y_1$ is treated as a fixed initial value which does not depend upon $X_1$. We denote the conditional density or probability mass function of $Y_t$ given $Y_{t-1} = y_{t-1}$ and $X_t=i$ as $f^{(i)}(\cdot | y_{t-1}; \theta^{(i)})$ or simply $f^{(i)}(\cdot | y_{t-1})$. This model is highly general, as $f^{(i)}(\cdot | y_{t-1})$ can be any valid density or probability mass function that depends upon the parameters $\theta^{(i)}$ and the previous observation $Y_{t-1}$. 
As a concrete example, if $Y_t$ is a scalar, then one may assume that $Y_t$ given $X_t = i$ is Normally distributed with parameters $\theta^{(i)} = \{\mu^{(i)},\sigma^{(i)},\phi^{(i)}\}$, where:

\begin{align}
\label{eqn:carhmm}
    \begin{split}
    \mathbb{E}(Y_{t}|Y_{t-1} = y_{t-1},X_t=i) &= \phi^{(i)} ~ y_{t-1} ~+ ~(1-\phi^{(i)})  ~\mu^{(i)}, \\
    \mathbb{V}(Y_t| Y_{t-1} = y_{t-1},X_t=i) &= (\sigma^{(i)})^2.
    \end{split}
\end{align}
%
A CarHMM which follows Equation (\ref{eqn:carhmm}) can be viewed as a discrete time version of a state-switching Ornstein-Uhlenbeck process \citep{Michelot:2019}. This follows in the same way that an AR(1) process is the discrete-time version of a traditional Ornstein-Uhlenbeck process.

As previously, the likelihood corresponding to a general CarHMM can be easily calculated using the forward algorithm. If $y$ is the sequence of observed emissions, then

\begin{equation*}
    \calL_{\text{CarHMM}}(\theta,\Gamma;y) = \delta \prod_{t=2}^T \Gamma P(y_t|y_{t-1};\theta) \mathbf{1}_N,
    \label{eqn:CarHMM_likelihood}
\end{equation*}
where
%
$P(y_t|y_{t-1};\theta)$ is an $N \times N$ diagonal matrix with $(i,i)^{th}$ entry equal to $f^{(i)}(y_t|y_{t-1}; \theta^{(i)})$. Figure \ref{fig:models}b shows a graphical representation of the dependence structure of a CarHMM.

\subsection{Incorporating multiple scales with the HHMM}
\label{subsec:HHMM}

A hierarchical hidden Markov model, or HHMM, accounts for processes occurring simultaneously at different scales by modelling both the coarse-scale process and fine-scale process with either HMMs \citep{Barajas:2017,Adam:2019} or CarHMMs. The coarse-scale model is either an HMM and CarHMM as defined in Sections \ref{subsec:HMM} and \ref{subsec:CarHMM}, where $\{X_1, \ldots, X_T\}$ is an unobserved Markov chain with $N$ possible states and $\{Y_1, \ldots, Y_T\}$ is the set of corresponding observations with state-dependent parameters $\theta^{(i)}$ for $i = 1,\ldots,N$.   
%
In the hierarchical setting, however, each state $X_t$ also emits another sequence of fine-scale unobserved states, $X^*_t \equiv \{X_{t,1}^*,\ldots, X_{t,T_t^*}\}$. If $X_t = i$, then $X^*_{t,t^*} \in \{1,\ldots,N^{*(i)}\}$ represents one of $N^{*(i)}$ possible within-curve behaviours associated with curve type $i$. In turn, $X^*_t$ emits a sequence of fine-scale observations $Y^*_t \equiv \{Y_{t,1}^*,\ldots, Y_{t,T_t^*}\}$. For each curve $t$, the fine-scale process $\{X^*_t, Y^*_t\}$ follows another HMM (or CarHMM) whose parameters depend on the value of $X_t$. If $X_t = i$, then the components of $X^*_t$ make up a Markov chain that has $N^{*(i)}$ possible states, an $N^{*(i)} \times N^{*(i)}$ transition probability matrix $\Gamma^{*(i)}$, and an initial distribution $\delta^{*(i)}$ which we assume is equal to the stationary distribution of the chain. The distribution of $Y^*_{t,t^*}$ given $Y^*_{t,t^*-1} = y^*_{t,t^*-1}$, $X^*_{t,t^*}=i^*$, and $X_t=i$ is governed by the parameter $\theta^{*(i,i^*)}$ and has density or probability mass function denoted $f^{*(i,i^*)}\left(\cdot|y^*_{t,t^*-1}; \theta^{*(i,i^*)}\right)$ or simply $f^{*(i,i^*)}(\cdot|y^*_{t,t^*-1})$. We denote the set of fine-scale emission parameters corresponding to $X_t=i$ as $\theta^{*(i)}=\big\{\theta^{*(i,1)}, \ldots, \theta^{*\left(i,N^{*(i)}\right)}\big\}$. In summary:

\begin{gather*}
    \{Y, X\} \text{ follows a (Car)HMM with } \Gamma \in \bbR^{N \times N}, \\
    %
    (Y_t | Y_{t-1} = y_{t-1}, X_t = i) \text{ has density } f^{(i)}(\cdot|y_{t-1};\theta^{(i)}), \\
    %
    \{Y^*_t,X^*_t | X_t = i\} \text{ follows a (Car)HMM with } \Gamma^{*(i)} \in \bbR^{N^{*(i)} \times N^{*(i)}}, \\
    %
    (Y^*_{t,t^*} | Y^*_{t,t^*-1} = y^*_{t,t^*-1}, X^*_{t,t^*} = i^*, X_t = i) \text{ has density } f^{*(i,i^*)}(\cdot|y^{*}_{t,t^*-1};\theta^{(i,i^*)}).
\end{gather*}
Given the coarse-scale hidden state sequence $X$, the $T+1$ sets $\{X_1^*, Y_1^*\}, \ldots, \{X^*_T, Y^*_T\}$, and $\{Y_1,\ldots,Y_T\}$ are assumed to be independent of one another.

Forcing certain parameters to be shared can reduce the complexity and increase the interpretability of an HHMM. For example, in our killer whale case study (see Section \ref{sec:data}), we take $N^{*(i)} = N^*$ for all $i$. We also share the fine-scale emission parameters across the $N$ coarse-scale hidden states $\left( \text{i.e., } \theta^{*(1,i^*)} = \cdots = \theta^{*(N,i^*)} = \theta^{*(\cdot,i^*)} \text{ for all } i^* = 1, \ldots, N^* \right)$. Coarse-scale hidden states therefore differ only in their coarse-scale emission parameters $\theta^{(i)}$ and fine-scale probability transition matrices $\Gamma^{*(i)}$, \textit{not} their fine-scale emission distributions.

While at first it seems difficult to jointly infer the paramtersDue to the nested structure of the HHMM, its likelihood is easily calculated using the forward algorithm.
%
Let $y$ be the sequence of observed coarse-scale emissions and
$y^* \equiv \{y^*_1, \ldots,y^*_T\}$ be the collection of $T$ observed fine-scale emission vectors.
%
In addition, let $\theta^* \equiv \{\theta^{*(1)}, \ldots, \theta^{*(N)}\}$ denote the collection of all fine-scale emission parameters and $\Gamma^* \equiv \{\Gamma^{*(1)}, \ldots, \Gamma^{*(N)}\}$ denote the collection of all fine-scale transition probability matrices. The likelihood of the observed data is then

\begin{equation}
    \calL_{\text{HHMM}}(\theta,\theta^*,\Gamma,\Gamma^*;y,y^*) = \delta P_m(y_1,y_1^*;\theta,\theta^*,\Gamma^*) \prod_{t=2}^T \Gamma P_m(y_t,y_t^*|y_{t-1};\theta,\theta^*,\Gamma^*) \mathbf{1}_N,
    \label{eqn:HHMM_likelihood}
\end{equation}
%
where $P_m(y_t,y_t^*|y_{t-1};\theta,\theta^*,\Gamma^*)$ is an $N \times N$ diagonal matrix whose exact structure depends upon the coarse- and fine-scale models. If the coarse-scale model is an HMM, $P_m(y_1,y_1^*;\theta,\theta^*,\Gamma^*)$ and $P_m(y_t,y_t^*|y_{t-1};\theta,\theta^*,\Gamma^*)$ for $t \geq 2$ both have $(i,i)^{th}$ entries equal to 
$f^{(i)}(y_t)\calL_{\text{fine}}\left(\theta^{*(i)},
\Gamma^{*(i)};y_t^*\right)$. 
If the coarse-scale model is a CarHMM, $P_m(y_1,y_1^*;\theta,\theta^*,\Gamma^*)$ has $(i,i)^{th}$ entry equal to $\calL_{\text{fine}}\left(\theta^{*(i)},
\Gamma^{*(i)};y_1^*\right)$ and $P_m(y_t,y_t^*|y_{t-1};\theta,\theta^*,\Gamma^*)$ for $t \geq 2$ has $(i,i)^{th}$ entry equal to $f^{(i)}(y_t|y_{t-1})\calL_{\text{fine}}\left(\theta^{*(i)},
\Gamma^{*(i)};y_t^*\right)$.
The fine-scale likelihood $\calL_{\text{fine}}$ corresponds to the likelihood of the fine-scale model, which can be either a CarHMM or an HMM. Figure \ref{fig:models}c graphically displays the dependence structure for an HHMM.

\subsection{Transforming fine-scale observations with the HMM-DFT}
\label{subsec:STFT}

In many applications where data are collected at high frequencies, intricate dependency structures arise within the fine-scale process that cannot be adequately modelled with the HMM-based models described thus far. To handle these additional fine-scale structures, we recommend replacing $Y^*_t = \{Y^*_{t,1},\ldots,Y^*_{t,T^*_t}\}$ with relevant statistics that summarize any non-Markovian behaviour. To maintain the temporal structure of the fine-scale process, local summary statistics can be calculated from a moving window with stride length $h$ over the elements of $Y^*_t$. Stride length refers to the distance between the first element of consecutive windows, so a stride length of $h$ implies that the first window starts at $Y^*_{t,1}$, the second at $Y^*_{t,1+h}$, and so on.
Replacing $Y^*_{t,t^*}$ with relevant summary statistics results in a loss of information since we are substituting infinite-dimensional curves with a low-dimensional representation. \citet{Aue:2015} and \citet{Gao:2019} use functional principle component analysis (FPCA) for dimension reduction to minimize error when predicting functional data. However, these papers do not account for state-switching behaviour between curves and do not guarantee that the resulting FPCA scores will be interpretable for practitioners. The goal of this paper is to classify and describe curve types and within-curve behaviour, not predict future functional data. We are thus primarily interested in the fine-scale summary statistics themselves rather than the curves from which they are calculated, and these summary statistics are known to have consistent maximum likelihood estimates \cite{Douc:2004}. As a result, we choose a dimension reduction technique which can effectively separate behavioural states and has an intuitive interpretation for practitioners.
In general, subject matter experts are often required to determine the specific summary statistics employed as well as the optimal window size and stride length of the moving window. Larger stride lengths result in a larger loss of information but also reduce the dimension of the fine-scale process, which allows for faster model fitting. In addition, setting the stride length equal to the window size avoids artificial residual correlation arising from overlapping windows. For our particular case study, we use the discrete Fourier transform (DFT) of a moving forward window with width $h$ and stride $h$ across $Y^*_t$:

\begin{align}
    DFT\{Y^*_{t,t^*},\ldots, Y^*_{t,t^*+h-1}\}(k) = \sum_{n=0}^{h-1} Y^*_{t,t^*+n}\exp\left(-\frac{i 2\pi}{h} kn \right)
    \label{eq:DFTdef}
\end{align}
%
for $t^* = 1,h+1,2h+1,\ldots$ and $k = 0, 1, \ldots, h-1$ with $i = \sqrt{-1}$. If $Y^*_{t,t^*}$ is a vector then the DFT is taken component-wise. We omit the final window if $t^*+h-1$ exceeds $T^*_t$, denote the total number of windows as $\tilde T^*_t$, and index each window with $\tilde{t}^* = 1,\ldots,\tilde T^*_t$. Next, we calculate transformed observations $\Z_{t,\tilde{t}^*} \equiv \{\Zone_{t,\tilde{t}^*},\Ztwo_{t,\tilde{t}^*}\}$:

\begin{equation}
    \label{eqn:z}
    \Zone_{t,\tilde{t}^*} \equiv \frac{1}{h}\sum_{n=1}^{h}Y^*_{t,h(\tilde{t}^*-1)+n}, \quad \Ztwo_{t,\tilde{t}^*} \equiv \sum_{k=1}^{\tilde{\omega}}\Big|\Big|DFT\{Y^*_{t,h(\tilde{t}^*-1)+1},\ldots, Y^*_{t,h\tilde{t}^*}\}(k)\Big|\Big|_2^2,
\end{equation}
%
where $\tilde{\omega} \leq h-1$ is a problem-specific tuning parameter corresponding to the maximum recorded frequency within each window. In words, $\Zone_{t,\tilde{t}^*}$ is the average value of $Y^*_t$ within window $\tilde{t}^*$ and $\Ztwo_{t,\tilde{t}^*}$ is the squared two-norm of the component of the window that can be attributed to frequencies between one and $\tilde{\omega}$ periods per window. More intuitively, $\Ztwo_{t,\tilde{t}^*}$ corresponds to the ``wiggliness" of the fine-scale data within curve $t$ and window $\tilde{t}^*$.

When performing this transformation, the fine-scale HMM (or CarHMM) must be redefined since $\Z_t$ exists on a coarser scale than $Y^*_t$ itself. As a result, there are only $\tilde{T}_t^*$ unobserved states associated with $\Z_t = \big\{\Z_{t,1},\ldots, \Z_{t,\tilde{T}^*_t}\big\}$, which we denote as $\tilde{X}^*_t = \big\{\tilde{X}^*_{t,1},\ldots, \tilde{X}^*_{t,\tilde{T}^*_t}\big\}$. The fine-scale probability transition matrices  $\Gamma^{*(i)}$ and probability density functions $f^{*(i,i^*)}$ are then applied directly to $\tilde{X}^*_t$ and $\Z_t$ instead of $X^*_t$ and $Y^*_t$. The likelihood of this model is therefore identical to that of the original HMM or CarHMM as defined in Sections \ref{subsec:HMM} and \ref{subsec:CarHMM}, but $Y^*$ is replaced with $\Z$ and $X^*$ is replaced with $\tilde{X^*}$. To clearly differentiate models, we refer to an HMM with $\Z$ as observations and $\tilde{X^*}$ as hidden states as an HMM-DFT. Figure \ref{fig:models}d displays the dependence structure of a fine-scale HMM-DFT.

\subsection{Generalized hierarchical Markov models}

Traditional HHMMs treat both the coarse-scale and the fine-scale processes as realizations of an HMM or CarHMM. 
However, the fine-scale observations of a particular dive $Y^*_t$ can be modelled using a large variety of parametric models which admit easy-to-compute likelihoods or penalized likelihoods. As such, the fine-scale HMM likelihood term $\calL_{\text{fine}}$ in Equation (\ref{eqn:HHMM_likelihood}) can be replaced by the likelihood of a general fine-scale model whose parameters depend upon the coarse-scale hidden state.
%
Possible candidates for the fine-scale model include any of the models described in the previous subsections in addition to many others not described here. 
For example, \citet{Bebbington:2007} and \citet{Borchers:2013} investigate data sets with count onsets as observations, so they use variations of a Poisson process as their fine-scale model. If the fine-scale model is a simple Poisson process, then this approach is equivalent to a Markov-modulated Poisson process \citep{Fischer:1993}.
The fine-scale process can also be modelled similarly to \citet{Langrock:2018}, who use B-splines to model the emission distribution of an HMM. This non-parametric approach uses a penalized likelihood term which can easily replace the usual fine-scale likelihood term in Equation (\ref{eqn:HHMM_likelihood}). 
Another class of fine-scale models is the set of continuous time methods such as a continuous-time HMM (CTHMM) \citep{Liu:2015} or a state-switching Ornstein-Uhlenbeck process \citep{Michelot:2019}. 
A continuous-time HMM may be appropriate if observations are not equi-spaced in time \citep{Liu:2015}. \citet{Xu:2018} model high-frequency biologging accelerometer data of individuals by incorporating a CTHMM into a hierarchical model similar to ours. However, they assume that individuals are partitioned into subgroups a priori whereas we use an HMM to infer the coarse-scale hidden states.

These examples include a few of many fine-scale models that can act as initial building blocks in a practitioner's toolbox to construct increasingly complex hierarchical models. 
A myriad of possible models can be built using this framework, but these models can quickly become complicated and computationally expensive to fit. Therefore, models should be constructed with care to achieve an adequate fit of the data while avoiding over-fitting and high computational costs.