### 2021-09-16 Report/main.tex

% Initial and revised submissions should be 12 point; this will be removed in the final version.
%\documentclass[12pt]{TD-CJS}
\documentclass{TD-CJS}

% Initial and revised submissions should also be double spaced. This command will be removed in the final version.
%\renewcommand{\baselinestretch}{2}

\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{psfrag}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{url}
%\usepackage{lineno}
\input{header}
\input{defs}

%\linenumbers

\begin{document}
\firstpage{1}
\lastpage{34}
\jvol{xx}
\issue{yy}
\jyear{2021}
\jid{CJS}
\aid{???}
% The running head contains the author names
\rhauthor{Sidrow et al.}
\copyrightline{Statistical Society of Canada}
\Frenchcopyrightline{Soci\'et\'e statistique du Canada}
% History: received and accepted dates
\received{\rec{14}{January}{2021}}
\accepted{\acc{14}{September}{2021}}

% User-defined commands go here
\renewcommand{\eqref}[1]{(\ref{#1})}
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mt}[1]{\mathrm{#1}}
\newcommand{\rv}{random variable}
\newcommand{\newblock}{}
\bibliographystyle{apalike}

% Title, authors, affiliations
%\title[]{Modelling temporal dependence and multilevel structure using hidden Markov models}%\query{Q1}
\title[]{Modelling multi-scale state-switching functional data with hidden Markov models}
\author{Evan Sidrow\authorref{*1}}
\author{Nancy Heckman\authorref{1}}
\author{Sarah M. E. Fortune\authorref{2}}
\author{Andrew W. Trites\authorref{3,4}}
\author{Ian Murphy\authorref{5}}
\author{Marie Auger-M\'eth\'e\authorref{1,4}}
\affiliation[1]{Department of Statistics, University of British Columbia, Vancouver, Canada}
\affiliation[2]{Marine Mammal Research Unit, University of British Columbia, Vancouver, Canada}
\affiliation[3]{Department of Zoology, University of British Columbia, Vancouver, Canada}
\affiliation[4]{Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada}
\affiliation[5]{Department of Biostatistics, University of Florida, Gainesville, USA}

% Abstract, keywords, and classification codes
\startabstract{%
\keywords{
\KWDtitle{Key words and phrases}
Accelerometer data\sep animal movement\sep biologging\sep diving behaviour\sep hierarchical modelling\sep killer whales\sep state-switching\sep statistical ecology\sep time series.
% MSC 2010 subject classification codes
\KWDtitle{MSC 2010}Primary 62M05 \sep secondary 62P12}
\begin{abstract}
\abstractsection{}
\abstractsection{Abstract}

Data sets comprised of sequences of curves sampled at high frequencies in time are increasingly common in practice, but they can exhibit complicated dependence structures that cannot be modelled using common methods of Functional Data Analysis (FDA). We detail a hierarchical approach which treats the curves as observations from a hidden Markov model (HMM). The distribution of each curve is then defined by another fine-scale model which may involve auto-regression and require data transformations using moving-window summary statistics or Fourier analysis. This approach is broadly applicable to sequences of curves exhibiting intricate dependence structures. As a case study, we use this framework to model the fine-scale kinematic movement of a northern resident killer whale ({\em{Orcinus orca}}) off the western coast of Canada. Through simulations, we show that our model produces more interpretable state estimation and more accurate parameter estimates compared to existing methods.

\abscopyright
\fabstractsection{}
\fabstractsection{R\'{e}sum\'{e}}
Ins\'{e}rer votre r\'{e}sum\'{e}
ici. We will supply a French abstract for those authors who
can't prepare it themselves.\Frenchabscopyright
\end{abstract}}
\makechaptertitle

% Email address for corresponding author
\correspondingauthor[*]{\\\email{evan.sidrow@stat.ubc.ca}} % only put this in after acceptance
%
\section{INTRODUCTION}
\input Sections/intro.tex
%
\section{MODELS AND PARAMETER ESTIMATION}
\label{sec:models}
\input Sections/models.tex
%
\section{KILLER WHALE CASE STUDY}
\label{sec:data}
\input Sections/results.tex
%
\section{SIMULATION STUDY}
\label{sec:sim_study}
\input Sections/simulation.tex
%
\section{DISCUSSION}
\input Sections/discussion.tex
%
\begin{ack}{ACKNOWLEDGEMENTS}
All killer whale data was collected under University of British Columbia Animal Care Permit no. A19-0053 and Fisheries and Oceans Canada Marine Mammal Scientific License for Whale Research no. XMMS 6 2019.
This project was supported in part by a financial contribution from Fisheries and Oceans Canada and the Natural Sciences and Engineering Research Council of Canada (Whale Science for Tomorrow).
This research was enabled in part by support provided by WestGrid (www.westgrid.ca) and Compute Canada (www.computecanada.ca).
We acknowledge the support of the Natural Sciences and Engineering Research Council of Canada (NSERC) as well as the support of Fisheries and Oceans Canada (DFO). 
Marie Auger-M\'eth\'e and Nancy Heckman thank the NSERC Discovery program, and Marie Auger-M\'eth\'e additionally thanks the Canadian Research Chair program.
Evan Sidrow thanks the University of British Columbia and the Four-Year Doctoral Fellowship program.
We are grateful to Dr. Joe Watson for his constructive suggestions.
\end{ack}
%
\newpage
\bibliography{references}
%
\newpage
\begin{appendix}
\input Sections/appendix.tex
\end{appendix}
%
\newpage
\section{FIGURES AND TABLES}
\input Sections/figures.tex
%
%\CJShistory

\end{document}